### 2023-06-30 paper.tex
feat: finish content

\documentclass{article}

\usepackage[title]{appendix}
\usepackage{bera} % not sure I love this font for this document
\usepackage{cite}
\usepackage{caption}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xcolor}

\lstset{
    basicstyle=\bfseries\ttfamily,
    breaklines=true
}

\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}

\lstdefinelanguage{json}{
    basicstyle=\normalfont\ttfamily,
    numbers=left,
    numberstyle=\scriptsize,
    stepnumber=1,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    frame=lines,
    backgroundcolor=\color{background},
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
}

\definecolor{lightgray}{rgb}{.9,.9,.9}
\definecolor{darkgray}{rgb}{.4,.4,.4}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
\lstdefinelanguage{OldTypeScript}{
  keywords={abstract, any, as, boolean, break, case, catch, class, console, const, continue, debugger, declare, default, delete, do, else, enum, export, extends, false, finally, for, from, function, get, if, implements, import, in, infer, instanceof, interface, keyof, let, module, namespace, never, new, null, number, object, package, private, protected, public, readonly, require, return, set, static, string, super, switch, symbol, this, throw, true, try, type, typeof, undefined, unique, unknown, var, void, while, with, yield},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  ndkeywords={class, export, boolean, throw, implements, import, this},
  keywordstyle=\color{blue}\bfseries,
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  sensitive=true,
  backgroundcolor=\color{lightgray},
  extendedchars=true,
  basicstyle=\ttfamily,
  showstringspaces=false,
  showspaces=false,
  numbers=left,
  numberstyle=,
  numbersep=9pt,
  tabsize=2,
  breaklines=true,
  showtabs=false,
  captionpos=b
}

\colorlet{comment}{green}
\lstdefinelanguage{TypeScript}{
    basicstyle=\normalfont\ttfamily,
    numbers=left,
    numberstyle=\scriptsize,
    stepnumber=1,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    frame=lines,
    backgroundcolor=\color{background},
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    commentstyle=\color{green}\ttfamily
}

\graphicspath{ {./images/} }

\title{Developing a Visual Studio Code Language Support Extension for the Snail Programming Language}
\author{Charles Reinhardt}

\begin{document}
\maketitle

\begin{abstract}

    A high-quality programming environment (often an integrated development environment, or IDE) can be vital to enhancing developer productivity. Visual Studio Code (VS Code) is a popular, open-source text editor maintained by Microsoft. VS Code delivers language-specific features through freely dowloadable, community-built extensions on an online marketplace. Many of these extensions allow developers to take advantage of editing features such as syntax highlighting, code autocompletion, or debugging support. The snail language (Strings Numbers Arrays and Inheritance Language) is a simple, object-oriented programming langauge meant to be implemented in a one-semester undergraduate course. We present a new VS Code extension to provide language support for the snail language. The extension implements support for syntax highlighting, rudimentary autocompletion, and static error-checking diagnostics using VS Code's Langauge Server Protocol. This report summarizes the contents of a VS Code extension and gives an overview of how an extension runs, particularly highlighting the functions of VS Code's Langauge Server Protocol. I also discuss how this extension can be futher developed to make use of VS Code's Debug Adapter Protocol to implement a debugger with breakpoints, start/stop behavior, and variable inspection.

\end{abstract}

\section{Introduction}

Much of software development in the present day takes place in integrated development environments (IDEs) \cite{JetBrains_2019}. An IDE is a collection of software development tools, such as a code editor, debugger, and build system, often unified under a similar user interface, with the goal of simplifying the software development process and enhancing developer productivity \cite{Gillis_Silverthorne_2018,Shyniaieva_2023}. In addition to composing these tools together, many IDEs also offer advanced features within their code editors such as syntax highlighting, code autocompletion, and error-checking diagnostics. Today, there are any number of different IDEs available for use with any given programming language, many offered as freely-downloadable for public use. For example, a developer looking to write code in Java may choose to do so in Eclipse, IntelliJ IDEA, or BlueJ \cite{EclipseFoundation_2023, JetBrains_2023,KingsProgrammingEducationToolsGroup_2022}. With so many high-quality IDEs available today, it is no surprise that 75\% of software developers today use an IDE in their everyday work \cite{JetBrains_2019}. Clearly, the IDE has become an integral part of the software development process today \cite{Vaniukov_2023}.

Visual Studio Code (VS Code), is a popular, open source text editor maintained by Microsoft \cite{StackOverflow_2022,Microsoft_2023a}. When first downloaded, VS Code is a lightweight text editor with minimal features. However, a number of community-built, freely downloadable extensions offer advanced langauge features. These extensions can be dowloaded on VS Code's online extension marketplace, and can turn VS Code into a very robust and powerful development environment for any programming task or language \cite{Microsoft_2023b}.

The snail programming language (Strings Numbers Arrays and Inheritance Language) is a simple, object-oriented programming language meant to be implemented in a one-semester undergraduate course \cite{Angstadt_2023a}. In order to be implemented in a short time frame, snail is defined by limited features and a relatively annoying syntax. For example, the language lacks a for-loop structure, opting instead to provide only while-loops. Each if-statment \emph{requires} an else clause, even when the developer does not need to use one. Every statement must end with a semi-colon, which is not a huge issue. Unless a developer forgets to include a line-ending semi-colon, where they will discover that the resulting parse error message is wildly uninformative. While this design makes snail easier to implement, it makes it hard for a developer to write programs in snail.

Currently, there are no tools to offer advanced langauge support for the snail language. This is no surprise, as snail has a small user base and was first released in February 2022 \cite{Angstadt_2023b}. With no external support for the language, software developers are taken out of their comfort zone and are offered no guidance when navigating the snail language.

This report presents the Snail Language Support VS Code extension, which seeks to address the lack of programming support tools for the snail programming language. Snail Langauge Support provides several important features to make programming in the snail language easier. First, it features syntax highlighting to make reading snail code easier and help highlight key structures or keywords of the language. Further, it features rudimentary autocompletion with auto-closing brackets, braces, and quotes, as well as if-else, while-loop, and class definition snippets, reducing the burden of memorizing snail's strict and unintuitive syntax. The extension also has automatic, real-time error checking diagnostics that allow a user to see syntax or parse errors in a snail program before running it for themselves. Finally, the extension is structured to support a full debugger with breakpoints, step-in, step-over, and step-out functionality.

This paper will outline the process of building the Snail Language Support VS Code extension.\footnote{This paper is formatted using \LaTeX. Recognizing that using \LaTeX \hspace{1pt} is really hard, we make the \LaTeX \hspace{1pt} source code for this paper available on GitHub: \url{https://github.com/CharlesReinhardt/snail-language-support-paper/}} Specificaly, we will introduce the structure of a VS Code extension that is meant to add support for a new programming langauge. We will also discuss how this extension uses VS Code's language server protocol (LSP) to provide real-time error diagnostics \cite{Microsoft_2022a}. We will address how the Snail Language Support extension may be further developed to include debugging support with breakpoints, step in/out behavior, and variable inspection, particularly highlighting the role of VS Code's debug adapter protocol (DAP) \cite{Microsoft_2021a}. Finally, we will review good software development practices such as version control and documentation.

\section{Background}

In this section, we will discuss the history of integrated development environments (IDEs) and debuggers, and how they both assist software developers today. We will also discuss Visual Studio Code (VS Code) in more detail, identifying the technologies that power VS Code.

\subsection{History of the Modern Integrated Development Environment}

The first programming environment to remotely resemble a modern IDE was the Dartmouth BASIC programming language run on the Dartmouth Time Sharing System (DTSS), developed in the mid 1960s \cite{KemenyKurtz_1968, Kurtz_1978}. Dartmouth BASIC was an example of a compile and go system, where program compilation was not separated from program execution \cite{Weik_2001}. Additionally, the DTSS also placed focus on making sharing time on a single university computer a simpler task, in order to help make programming more accessible to novices. This makes the DTSS an early example of combining multiple software development tasks into a single programming environment.

Borland's TurboPascal takes this idea one step further. Released in 1983, TurboPascal featured a Pascal compiler, code editor, file navigation user interface, and a rudimentary debugger \cite{Gajic_2017, Intersimone_2010, Wadlow_1984}. This time period would also see language-specific IDEs in Microsoft's Visual BASIC 1.0, Microsoft's QuickC, and Borland's Turbo C/C++, all featuring similar characteristics of early IDEs \cite{Burgwin_2013, Plant_Murrell_2007, Posch_2023, Forer_Petreley_1991}.

While early IDEs certainly had their merits, they usually only supported one programming language. Microsoft's Visual Studio, released in 1997, was one of the first IDEs to package support for a variety of programming languages in one piece of software \cite{Microsoft_1997}. Visual Studio also featured extensive tools to aid in development for software on the early internet.

Today, we see a trend towards open source IDEs such as Eclipse, NetBeans, or VS Code \cite{Apache_2023, EclipseFoundation_2023, Microsoft_2023a}. Even open source editors such as these include intelligent code editors, debugging tools, and version control integration. Many of these editors have advanced features within their code editors, providing functions such syntax highlighting, code autocompletion, and real-time error diagnostics.

There are many benefits to developing software using modern IDEs. By bundling code editing, build systems, and program execution into one tool, modern IDEs reduce the time and effort a developer needs to put forth in order to write and test a piece of code \cite{Gillis_Silverthorne_2018}. This also reduces the number of decisions a developer has to make while developing code, which can help increase productivity. Using an IDE can also standardize the software development process, by either helping a group of people use a consistent UI (and know how to help eachother), or allow a developer to avoid switching applications to complete a single task \cite{Veracode_2020}.

\subsection{History of Modern Debugging Tools}

Debugging is the process of searching for and fixing unexpected errors in a piece of code \cite{AWS_2023}. Early techniques of debugging software involved physically printing machine output and reading through it step-by-step, searching for a potential error \cite{RevDeBug_2020}. Some computer systems, such as the IBM 704 Data Processing system released in 1968, allowed programmers to print information stored in specified memory locations in specified forms to assist with debugging activities \cite{MIT_1957, IBMArchives_2023}.

The \lstinline{adb} and \lstinline{dbx} debuggers for the Unix operating system saw a release in the 1970s and 1980s \cite{BellLaboratories_1983, Linton_1990}. These tols introduced the concept of breakpoints, which help a developer pinpoint the location of a fault in a piece of code. Both \lstinline{adb} and \lstinline{dbx} were run on the command line. Next, came the GNU project debugger (\lstinline{gdb}), another command line debugger, released in 1986, which gave the user even greater control of tracing a program's execution throughout its runtime \cite{GNU_FSF_2023}. With the ability to debug low-level languages like C and Assembly, \lstinline{gdb} is still in use today.

Modern debuggers with graphical user interfaces (GUIs) include the Visual Studio Debugger, which allows a developer to track variable values, function calls, and even change pieces of code or values of expressions while debugging \cite{Microsoft_2022b}. Eclipse, a popular IDE for Java, has a similar debugger with a user friendly GUI \cite{Sinha_2017}. Today, many developement workflows include the use of debugging tools \cite{Matloff_Salzman_2008}.

Debugging tools can be of great assistance to developers. By allowing developers to more closely and quickly inspect a program's execution, debugging tools can reduce the amount of time a developer spends debugging, and thus enhance developer productivity \cite{ZhRaAmRuWo_2008}.

\subsection{What is Visual Studio Code?}

Visual Studio Code (VS Code) is a popular, open source code editor. VS Code is designed to be fast and lightweight, with a focus on allowing a developer to write, test, and debug source code quickly \cite{Microsoft_2023c}. To achieve this, VS Code uses native, web, and language-specific technologies. Tools like Electron allow VS Code to run on multiple platforms and operating systems using common web technologies like JavaScript, HTML, and CSS \cite{Electron_2023}.

While VS Code is lightweight and fast, it still supports advanced features such as build or debugger tools available through the online VS Code extension marketplace \cite{Microsoft_2023b}. VS Code is designed for extensibility, providing robust APIs to allow independent developers to create and publish extensions \cite{Microsoft_2023d}. As a result, VS Code is able to grow and develop along with its community of users.

Existing extensions make VS Code a popular code editor \cite{StackOverflow_2022}. The CodeSnap extension allows a user to capture and share improved screenshots of code \cite{adpyke_2021}. GitLens adds additional visualization tools to help developers contribute to Git repositories while editing in VS Code \cite{GitKraken_2023}. Language Support for Java(TM) provides intellisense, formatting, refactoring, and build system support for the Java language \cite{RedHat_2023}. The result of this extension support is an IDE that is used by over 75\% of professional and hobby software developers \cite{StackOverflow_2022}.

\section{Anatomy of a Visual Studio Code Extension}

In this section, we will discuss the contents of a Visual Studio Code (VS Code) extension. First, we will highlight the directory structure of a typical language support VS Code extension. Next, we will discuss how a VS Code extension runs, from the moment it starts up to the moment it shuts down. We will also overview what can be provided to the user by a VS Code extension, specifically highlighting language configurations, autocompletion, and syntax highlighting.

\subsection{Directory Structure}

Most VS Code extensions share a similar directory structure. These structures may vary depending on what purpose the extension is meant to serve. In figure \ref{fig:directory-structure}, we see a diagram of a typical language support VS Code extension.\footnote{The exact directory structure of the Snail Language Support extension is not included in this paper. For a deeper look at the directory structure and source code for Snail Language Support, see the git repository hosted on GitHub: \url{https://github.com/snail-language/snail-language-support}} First, the \lstinline{.vscode} directory configures how we run and test our exstension locally. Through the \lstinline{launch.json} and \lstinline{tasks.json} files, we define various tasks and configurations to streamline the build and testing process.\footnote{From personal experience, we strongly recommend taking the time to understand these files before developing a VS Code extension. It will likely save painstaking debugging in the near future. See launch configurations: \url{https://code.visualstudio.com/docs/editor/debugging\#_launch-configurations} and tasks: \url{https://code.visualstudio.com/docs/editor/tasks}} Note that these configuration files do not get packaged with a released VS Code extension. Next, a \lstinline{README.md} file allows us to advertise the features of our extension. This file is rendered in the VS Code extension marketplace and is meant to be a user's first impression of an extension.

\begin{wrapfigure}{R}{0.3\textwidth}
    \begin{center}
        \includegraphics[height=0.3\textheight]{png/directory-structure.png}
        \caption{
            \centering 
                The typical directory structure of a VSCode language support extension.
        }
        \label{fig:directory-structure}
    \end{center}
\end{wrapfigure}

Snail Langauge Support is built with TypeScript. TypeScript is a syntactic superset of JavaScript that allows the developer the benefits of a type system without sacrificing JavaScript's flexibility \cite{TypeScript_2023}. TypeScript files are denoted with a \lstinline{.ts} suffix. Using the TypeScript compiler (\lstinline{tsc}), a program in TypeScript is transpiled to an equivalent JavaScript program. With a \lstinline{tsconfig.json} file in an extension directory, we define how the TypeScript compiler handles an extension's TypeScript code. We define where the compiler looks for our \lstinline{.ts} files, where to place resulting JavaScript files, and the strictness of type checks during transpilation.

The \lstinline{src} directory contains the aforementioned TypeScript source code that runs our extension. In this directory, the \lstinline{extension.ts} file defines special actions for our extension and launches the extension within an existing VS Code window. The \lstinline{server.ts} file defines a language server that allows our extension to take advantage of real-time error diagnostics. We further discuss language servers later in this paper.

Lastly, the \lstinline{package.json} file defines which features and configurations our extension supports. It also contains some biographical information about the extension, such as the author, publisher, or name of the exstension.

To build the structure for Snail Language Support, we used the \lstinline{Yeoman} tool to generate an extension skeleton \cite{Yeoman_2023}. This skeleton was instrumental in providing basic configuration and source files to modify into the Snail Language Support extension.

\subsection{Extension Lifecycle: From Startup to Shutdown}

When a VS Code window launches, it also launches all extensions the user has downloaded and enabled by calling each extension's \lstinline{activate} function, defined in the extension's source code. In Snail Language Support's case, the \lstinline{activate} function is defined in the \lstinline{extension.ts} file. This file has access to the VS Code extension API, allowing a developer to define certain VS Code client UI updates \cite{Microsoft_2023d}. For example, Snail Language Support can display a VS Code style error message to the user if the user does not have a compatibale version of the snail language dowloaded.

The \lstinline{activate} function is responsible for a few different behaviors. First, the \lstinline{activate} function registers multiple command handlers to watch for commands input from a user. A command handler is an example of the observer design pattern, which allows a developer to address a variety of potential input events that are experienced during runtime \cite{GangOfFour_1995}. A developer can register an observer that listens for a particular event and executes a block of code once that event is detected. Snail Language Support defines command handlers to detect when a user chooses to run or debug a snail file. We will touch on the topic of debugging within Snail Language Support more later.

The \lstinline{activate} function also launches a snail language server, which communicates with VS Code via VS Code's Language Server Protocol (LSP). We further discuss the topic of language servers later in this paper.

Once the extension setup inside the \lstinline{activate} function is complete, VS Code enables the extension within the workspace, allowing a developer to utilize the functions of the extension. Once the VS Code window is closed, or the extension is manually disabled, VS Code calls the \lstinline{deactivate} function, which shuts down any separate processes left over from the extension run.

To inspect Snail Language Support's \lstinline{extension.ts} file and \lstinline{activate} function more closely, see appendix \ref{app:extension-ts}

\subsection{Extension Contributions: The Extension Manifest}

Recall that the \lstinline{package.json} defines what our extension contributes to the user. We now discuss some of those contributions here.

First, we can define some metadata for our extension, such as a name, author, description, and version. We can also define which version of the VS Code engine this extension expects. This ensures that an extension knows which VS Code APIs it has proper access to. For example, an extension that expects a VS Code version of \lstinline{1.7.0} won't be able to take advantage of an API released in \lstinline{1.8.0} \cite{Microsoft_2023e}. We can also define some categories that our extension falls under, helping our extension gain visibility on the online VS Code marketplace.

Next, and very importantly, we define where the source code is located for our extension. Through the \lstinline{main} attribute, we define the entry point for our extension. This is the file that contains the \lstinline{activate} function. For Snail Language Support, this is the \lstinline{/out/extension.js} file. Notice that this is the transpiled version of our \lstinline{src/extension.ts} file. For a more detailed look at Snail Language Support's \lstinline{package.json} file, see appendix \ref{app:package-json}.

\subsubsection{Language Configuration}

Snail Language Support also provides language configuration settings supported by a VS Code extension. These configurations are found in \lstinline{language-configuration.json} file. We can define what is considered a comment through the \lstinline{comments.lineComment} and \lstinline{comments.blockComment}attributes, which allows a user to comment lines using VS Code's comment shortcut. Similarly, we can also define auto-closing structures such as parentheses or brackets through the \lstinline{autoClosingPairs} attribute, which allows VS Code to automatically place closing parentheses or brackets. This is particularly useful for snail, as most opening parentheses or brackets require a closing partner. By defining \lstinline{autoClosingPairs}, a developer can expend less mental energy ensuring that each parenthesis and bracket has a closing pair, and more mental energy on developing code. The \lstinline{language-configuration.json} file for Snail Language Support can be found in appendix \ref{app:language-configuration}. For more attributes and configuration options, see the VS Code documentation on language configurations \cite{Microsoft_2023f}.

\subsubsection{Rudimentary Autocompletion Through Snippets}

Snail Language Support provides autocompletion suppport for common structures such as if-else clauses, while-loops, and class definitions. This autocompletion is implemented through VS Code snippets. A snippet is a commonly used code structure that is defined by a \lstinline{prefix} (or a few options for a \lstinline{prefix}) and a \lstinline{body}. When a user types a snippet \lstinline{prefix}, they are given the option to replace the keyword with the \lstinline{body} of the snippet to complete the structure. Snippet bodies can also contain placeholders, which a user can visit in sequence while pressing their \lstinline{tab} button. For a full listing of the snippets that Snail Language Support provides, see appendix \ref{app:snippets-json}. Similarly, for more information on snippets in VS Code, see the documentation \cite{Microsoft_2023g}.

\subsubsection{Syntax Highlighting}

Snail Language Support also supports syntax highlighting for the snail language as shown in figure \ref{fig:syntax-highlight}. VS Code's syntax highlighting makes use of TextMate language grammars. A generic language grammar is a set of rules that govern what constitutes a valid statement for a given programming language \cite{pgrandinetti_2019}. A TextMate grammar is a specific format for defining a language grammar \cite{MacroMates_2021a}. TextMate grammars define patterns that match special elements of a text document and assign scope names to these elements. For example, keywords of a particular language might be assigned the \lstinline{keyword.control} scope name. These patterns are defined with regular expressions.

\begin{figure}[b]
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{screenshots/dev-no-highlight-light.png}
            \caption{Developing snail code without syntax highlighting can be a mentally taxing experience.}
            \label{fig:dev-no-highlight}
        \end{center}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{screenshots/dev-with-highlight-light.png}
            \caption{Syntax highlighting makes developing snail code much easier.}
            \label{fig:dev-with-highlight}
        \end{center}
    \end{subfigure}
    \caption{Demonatrating the benefit of Snail Language Support's syntax highlighting feature.}
    \label{fig:syntax-highlight}
\end{figure}

A regular expression is a string of text that can be used to match a particular pattern in a piece of text. Regular expressions are used in a variety of programming activities, such as lexing a program or to find text in a code editor \cite{ComputerHope_2022, Cornell_2022}.

TextMate grammars use the advanced Oniguruma regular expression syntax, developed by K. Kosako \cite{Microsoft_2023h, MacroMates_2021b}. The Oniguruma regular expression syntax offers advanced features such as case insentivity options, character groupings, quantifiers, and anchors.

Snail Language Support defines a TextMate grammar in the \lstinline{snail.tmLanguage.json} file, shown in appendix \ref{app:snail-tmLanguage}. Taking a closer look at the structure of this document, we see three main relevant attributes: \lstinline{$schema}, \lstinline{patterns}, and \lstinline{repository}.

First, the \lstinline{$schema} attribute links to a TextMate grammar schema document. This allows our editor to verify our \lstinline{snail.tmLanguage.json} file and ensure we are defining valid attributes.

Next, we define a series of \lstinline{patterns} using regular expressions.  VS Code's tokenization engine reads our snail file one line at a time and tries to match the text to a pattern (or combination of patterns) defined in our \lstinline{patterns} attribute. If the engine is successful in matching snail text to a pattern, it assigns that text a scope name. VS Code then uses these scope names to highlight the text appropriately, depending on the chosen color theme.

The engine tries matching text in the order we define our patterns. This means that the first pattern we define in \lstinline{patterns} is the first regular expression that the tokenization engine tries to match to a line in a snail file. It is important to pay attention to the order that patterns are defined. For example, Snail Language Support defines comments first, which ensures that all comments in snail, no matter how complicated or syntactically complex, is highlighted as a simple comment.\footnote{While defining our regular expression patterns, we found the Rubular program essential. It allowed us to iterate and test regular expressions more rapidly. We strongly recommend investigating this site while developing a TextMate grammar of your own: \url{https://rubular.com/}}

TextMate grammars also allow you to nest patterns inside of one another. This is essential for snail, as it allows the tokenization engine to match text inside of parentheses or brackets. TextMate grammars' ability to recursively reference the grammar itself (through the \lstinline{$self} keyword) proves essential for this functioanlity.

For Snail Language Support, we define our patterns in the \lstinline{repository} attribute. This allows us to reference them in multiple places in our \lstinline{patterns} attribute. We feel this makes the document easier to understand and modify when necessary.

It is important that Snail Language Support's syntax highlighting can handle a snail file with incorrect syntax. For example, if a snail file uses a class definition with the incorrect syntax, Snail Language Support should still be able to highlight the code contained in that class. To achieve this, we must make sure that all patterns are accessible from the top level of the grammar. While any pattern may be nested inside of another pattern, that pattern must be present elsewhere in the grammar so that it is not \emph{only} accessible after matching the other pattern.

\section{Language Servers and the Language Server Protocol (LSP)}

Snail Language Support also provides real-time lexing and parsing error diagnostics when developing code in snail. To do this, Snail Language Support uses a language server that communicates via VS Code's langauge server protocol (LSP). We now discuss what a language server is, how it relates to the LSP, and how Snail Language Support uses langauge servers.

\subsection{What is a Language Server?}

\begin{wrapfigure}{r}{0.5\textwidth}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{png/language-server.png}
        \caption{
            \centering 
            A development tool will communicate with a language server to display advanced diagnostics.
        }
        \label{fig:language-server}
    \end{center}
\end{wrapfigure}

A language server is a tool used to provide language-specific editing features such as advanced autocomplete, go-to definition, or automatic error checking \cite{Microsoft_2022a}. Many IDEs today utilize language servers by launching a language server process that runs in the background while a developer writes code. The development tool then sends \emph{requests} to the language server and the language server \emph{responds} with the information requested by the development tool. The development tool then displays the information to the developer. This process is displayed in figure \ref{fig:language-server}.

Many language servers are implemented in the language they emulate \cite{Microsoft_2023j}. For example, a typical Python language server would be implemented in Python. This makes integrating language servers into development tools difficult. First, most development tools for specific languages are not implemented in those specific languages. For example, most development tools for Python are not implemented with Python. Further, a single language server might have to interact with a number of different development tools. For example, a Python language server might have to interact with VS Code, PyCharm, and Spyder. Without a standard framework for communication between development tools and language servers, development tools and language servers must implement specific functionality for each pairing. This problem is displayed in figure \ref{fig:no-lsp-theory}.

Each development tool has different standards for how they interact with language servers. To address this, a development tool may need to adjust how they request and recieve information from each language server it interacts with. Alternatively, a language server might adjust how it recieves and responds to requests for each development tool it interacts with. Both options are not ideal.

\subsection{The Language Server Protocol}

The Langauge Server Protocol (LSP) defines a standard framework for development tool and language server communication \cite{Microsoft_2022a}. A development tool is able to communicate with any language-specific language server by sending requests and recieving responses according to the LSP standard.

The LSP defines a number of common interactions between a development tool and a langauge server. For example, a development tool might send the \lstinline{DidchangeTextDocument} notification when a development tool detects that a text document has been edited. A language server might validate this text document and send any notable diagnostics back to the development tool.

VS Code's LSP relies on requests and responses \cite{Microsoft_2023i}. A request is a message that a development tool sends to a language server, asking the language server to perform some operation. A language server responds to the request by sending a corresponding response message with information about the executed operation. In order to follow the LSP, a language server \textbf{must} send a corresponding response upon recieving any requests. For example, a development tool might send the \lstinline{Initialize} request to a language server, and a language server must respond with the \lstinline{InitializeResponse} before the development tool is allowed to send any further requests. This contract goes both ways: when a language server sends a request to a developement tool, the development tool must respond.

VS Code's LSP also uses notifications. A notification is a message that does not require a response. A development tool can send a notification to a language server and the language server can process the notification as it sees fit. Often, the language server will send information back to the development tool (in the form of another notification), but it is important to note that the language server is not required to respond.

With an established standard for communication as shown in \ref{fig:lsp-theory}, both development tools and language servers can streamline their communications systems. Additionally, a developer producing one of these tools or features can rely on a consistent standard to make the development process easier.

\begin{figure}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{png/no-lsp-theory.png}
            \caption{Without the Language Server Protocol, development tool and language server communication can get messy.}
            \label{fig:no-lsp-theory}
        \end{center}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{png/lsp-theory.png}
            \caption{The Language Server Protocol streamlines the communication between development tools and language servers.}
            \label{fig:lsp-theory}
        \end{center}
    \end{subfigure}
    \caption{Demonatrating the benefit of VS Code's Language Server Protocol.}
    \label{fig:lsp}
\end{figure}

In VS Code extensions, a language server (using LSP) can provide a number of different features for a programming language. By implementing support for the \lstinline{textDocument/hover} request, an exstension could display additional documentation about code when a user hovers their cursor over some text. By implementing support for the \lstinline{widnow/showMessage} request, an extension can display an error message in VS Code's UI when code fails to build. With the \lstinline{textDocument/publishDiagnostics} notification, an extension can validate that a program will run before a user actually tries to run it. For a full list of LSP specificatons, see the documentation \cite{Microsoft_2023i}.

\subsection{Language Servers in Snail Language Support}

In the Snail Language Support extension, we use a language server to validate snail files before runtime, checking for lexer and parser errors. With the extension enabled, VS Code is able to display lexer and parser errors, and highlight the relevant location of the error, before a developer runs the program themselves. This behavior is illustrated in figure \ref{fig:dev-diagnostics}.

\begin{figure}[t]
    \begin{subfigure}{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{screenshots/dev-no-diagnostics-light.png}
            \subcaption{
                \centering
                    Developing snail code without real-time error diagnostics can make it difficult to spot syntax errors.
                }
            \label{fig:dev-no-diagnostics}
        \end{center}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{screenshots/dev-with-diagnostics-light.png}
            \subcaption{
                \centering
                    With Snail Language Support, a developer can spot easily spot syntax errors and fix them before runtime.
                }
            \label{fig:dev-with-diagnostics}
        \end{center}
    \end{subfigure}
    \caption{
        \centering
            Demonstrating the benefit of real-time error diagnostics provided in Snail Language Support.
        }
    \label{fig:dev-diagnostics}
\end{figure}

To do this, Snail Language Support uses a snail language server. The majority of this language server is implemented in TypeScript in the \lstinline{server.ts} file, which can be found in appendix \ref{app:server-ts}. This file responds to connections from a VS Code client and handles a few important LSP requests such as \lstinline{Initialize} events and \lstinline{DidChangeTextDocument} notifications.

The snail-specific features are provided by modifying the snail interpreter to act as a language server. By passing the \lstinline{-s} or \lstinline{--server} flag while executing a snail file (i.e. \lstinline{snail -s hello-world.sl}), the snail interpreter outputs relevant error diagnostics including program exit status, error source, and error location.\footnote{Further modifying the snail interpreter to handle a variety of potential language server requests (not just requests for error diagnostics) is a great area for future work.} The \lstinline{server.ts} file handles this output and provides it to the VS Code client in the LSP format.

\begin{wrapfigure}{R}{0.5\textwidth}
    \includegraphics[width=0.5\textwidth]{png/lsp-sequence.png}
    \caption{
        \centering
            Modelling the interactions between a developer, VS Code client, and snail language server while a developer modifies a snail file.
    }
    \label{fig:lsp-sequence}
\end{wrapfigure}

A high-level sequence of interactions between a developer, VS Code client, and snail language server is documented in figure \ref{fig:lsp-sequence}. First, a user opens a program written in snail (1). When the VS Code client detects this activity, it launches a snail language server (2). Next, the VS Code client provides the snail program to verify (3) and the language server executes the snail program, using the aforementioned \lstinline{snail -s} option (4). The snail language server returns the error diagnostics to the VS Code client (5) and the client displays these diagnostics to the developer (6). While the developer further modifies the snail file, the VS Code client and snail language server continue to communicate via LSP (7-10). When a developer is finished, they can close the snail file (11) and the VS Code client shuts down the snail language server (12).

\section{Debug Adapter Protocol}

Snail Language Support is also structured to support the implementation of a debugger for snail. To do this, Snail Language Support defines a debug adapter that uses VS Code's Debug Adapter Protocol (DAP). In this section, we will discuss what the DAP is and how Snail Language Support uses it.

\subsection{What is the Debug Adapter Protocol?}

Debugging tools are an integral part of the software development process. Modern debuggers are often implemented with a user-friendly UI to help developers debug their programs \cite{Microsoft_2022b,Matloff_Salzman_2008,Sinha_2017}. Many of these debuggers are built intertwined with their corresponding code editor's UI, which can vary from code editor to code editor. As a result, valuable debugging logic is intermingled with editor-specific UI logic. This makes it nearly impossible to implement a debugger that is capable of communicating with different code editors \cite{Microsoft_2021a}. This problem is illustrated in figure \ref{fig:no-dap-theory}.

Instead of asking a development tool to interface directly with a debugger, suppose a development tool interfaces with a debug adapter that translates a development tool's requests and sends them to a language-specific (but tool-agnostic) debugger. If the communication between the development tool and debug adapter is standardized, language-specific debuggers and debug adapters can be reused across multiple development tools.

VS Code's Debug Adapter Protocol (DAP) standardizes the communication between a development tool and debug adapter. As shown in figure \ref{fig:dap-theory}, this allows language-specific debuggers and debug adapters to be reused for multiple development tools.

\begin{figure}[h]
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \includegraphics[width=\textwidth]{png/no-dap-theory.png}
        \caption{
            \centering
                Communication between a generic debugger and a development tool is incredibly complicated.
        }
        \label{fig:no-dap-theory}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \includegraphics[width=\textwidth]{png/dap-theory.png}
        \caption{
            \centering
                With the Debug Adapter Protocol, a development tool can easily communicate with a generic langauge-specific debugger.
        }
        \label{fig:dap-theory}
    \end{subfigure}
    \caption{
        \centering
            Illustrating the benefit of the Debug Adapter Protocol in aiding development tool and debugger communication.
    }
    \label{fig:dap}
\end{figure}

VS Code's DAP relies on requests and responses very similarly to the LSP \cite{Microsoft_2021b}. A request is a message that a development tool sends to a debug adapter, asking the debugger to perform some debugging task. The debug adapter communicates with the language-specific debugger to perform this operation and sends a corresponding response message back to the development tool. The debug adapter \textbf{must} send a corresponding response after recieving a request, even if the response is an empty message. The DAP also defines reverse requests, which function the same as regular requests except that they are sent from the debug adapter to the development tool.

VS Code's DAP also allows a debug adapter to send events to the development tool. This allows the debug adapter to update the development tool on a debugging session's state. For example, a debug adapter can send the development tool a \lstinline{Stopped} event when program execution has stopped due to a breakpoint, step request completion, or expression evaluation completion.

\subsection{Debug Adapter Protocol in Snail Language Support}

Snail Language Support does not currently provide a debugger for the snail language. However, the extension is structured to support a snail debugger in the future.

\begin{wrapfigure}{r}{0.5\textwidth}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{png/dap-highlevel.png}
        \caption{
            \centering 
            A high-level overview of VS Code, Snail Language Support, and snail debugging communication
        }
        \label{fig:dap-high-level}
    \end{center}
\end{wrapfigure}

Snail Language Support provides a debug adapter for the snail language in two parts, as shown in figure \ref{fig:dap-high-level}. First, we define an extension-level debug adapter within Snail Language Support in the \lstinline{debugAdapter.ts} file, as found in appendix \ref{app:debug-adapter-ts}. This debug adapter is launched using the \lstinline{node.js} runtime within the VS Code client. The VS Code client sends DAP requests to the extension-level debug adapter via standard input (\lstinline{stdin}), and the debug adapter uses the transmission control protocol (TCP) to forward these DAP requests along to snail. Think of the extension-level debug adapter as a tool to echo DAP requests from VS Code along to snail.

Next, the DAP requests are processed in a snail-level debug adapter that we define as an additional module alongside the snail interpreter in the \lstinline{snaildap.re} file as shown in appendix \ref{app:snaildap}. The snail-level debug adapter is responsible for responding to DAP requests with the necessary snail-specific information.

\subsubsection{The Snail-Level Debug Adapter}

We implement the snail-level debug adapter (\lstinline{snaildap}) using the Reason programming language. Reason is built on top of the functional and strictly-typed programming language, OCaml\cite{Reason_2020}. Reason introduces a JavaScript-like syntax on top of OCaml's type system, giving developers a familiar syntax while providing improved code safety and maintanability. Anything written in Reason can be translated to OCaml (and vice versa), allowing developers to take advantage of OCaml tools when developing in Reason.

The \lstinline{snaildap} debug adapter makes use of the \lstinline{ocaml-dap} library developed by Hackwaly \cite{Hackwaly_2021}. The \lstinline{ocaml-dap} library is an implementation for VS Code's DAP in the OCaml langauge, allowing a developer to reference relevant types, requests, and commands when handling DAP requests in OCaml. While using the \lstinline{ocaml-dap} library in Reason proves to be convenient, it does introduce one problem. After performing debug initialization, \lstinline{ocaml-dap} requires a program to, essentially, sit and wait for DAP requests to be recieved. This prevents \lstinline{snaildap} from actually running snail and aqcuiring program details to respond to DAP requests once they are recieved.

We solve this problem by introducing the Lightweight Thread (\lstinline{Lwt}) library for OCaml and Reason. The \lstinline{Lwt} library introduces the concept of a promise, a placeholder return object for time-consuming operations \cite{Ocsigen_2023}. When a developer wants a program to continue running while waiting for operation to finish, they can introduce a promise that contains an empty value. Once the time-consuming operation is finished, the promise becomes fulfilled and gains a meaningful value. During that time spent waiting for a promise to be fulfilled, a program is able to perform other operations. In the context of \lstinline{snaildap}, the \lstinline{Lwt} module allows \lstinline{snaildap} to perform snail interpreter operations \emph{and} wait for DAP requests \emph{simultaneously}.

Taking a look at the \lstinline{snaildap.re} file as shown in appendix \ref{app:snaildap}, the important initialization of our debug adapter can be found in three functions: the \lstinline{uninitialized}, \lstinline{initialized}, and \lstinline{debug} functions.

In the \lstinline{unitialized} and \lstinline{initialized} functions, \lstinline{snaildap} handles \lstinline{Initialize} and \lstinline{Launch} requests. If \lstinline{snaildap} were to also support the \lstinline{Attach} request, that behavior would be implemented here. The \lstinline{unitialized} and \lstinline{initialized} functions also handles \lstinline{Disconnect} requests when necessary.

Finally, in the \lstinline{debug} function, we define behavior for the DAP capabilities that our debug adapter supports. Currently, Snail Language Support handles \lstinline{ConfigurationDone}, \lstinline{Threads}, \lstinline{Terminate}, and \lstinline{Disconnect} requests. This is enough for Snail Language Support to start and stop a debugging session via the VS Code UI, but not enough for a user to do any debugging to a snail program. Further work to include real debugging logic would be to implement behavior for a variety of DAP requests as shown in appendix \ref{app:further-dap}.

\section{Good Practices in Software Development}

Software development is a growing industry where programmers work in teams to develop large, cohesive software to perform a variety of tasks, ranging from anywhere from gaming, healthcare, social media, and everything in between \cite{BusinessResearchCompany_2023}. With high-quality software in high demand, software developing teams must know how to work together efficiently. There are a number of tools that software developers use to help produce high-quality software. Two tools that the Snail Language Support extension emphasizes are version control and documentation.

\subsection{Software Version Control}

Software version control is a powerful tool in the software development world. In this section, we will first discuss why version control can be so valuable, and discuss a few different types of version control systems. We will discuss the git version control system how Snail Language Support uses of git.

\subsubsection{Introduction to Version Control Systems: Why Version Control?}

A large challenge in developing software in teams is creating one cohesive software product when developement is done simultaneously by different programmers on different machines. Version control is the task of keeping a software consistent, maintainable, and stable, despite containing many different versions and configurations undergoing development in parallel. Many different version control \emph{systems} exist for developers today to automate version control tasks for a team. There are two primary types of version control systems: centralized version control systems (such as Apache Subversion) and distributed version control systems (such as git) \cite{Collins-SussmanFitzpatrickPilato_2004, ChaconStraub_2014}.

In a centralized version control system, source code history is stored in a single, central server as shown in figure \ref{fig:centralized-vcs}. This central server is the "correct" version of a code repository. A developer working on code stored in a centralized version control system must get individual files from the server, make changes, and check those changed files directly back to the central server.

Centralized version control systems have the benefit of being straightforward and easy for developers to understand. Additionally, having only one central server for a code repository leaves no confusion as to where the software source code lies.

Centralized version control systems also come with their downsides. First, a developer must have an internet connection in order to make and test changes to a software. This is unfortunate for a developer who may like to do work when internet might not be available. Additionally, there are security risks associated with only maintaining one server of source code history. If the single server becomes vulnerable to attack, or if the server crashes for some reason, a software's entire code repository can be lost.\footnote{When first writing this paper, we neglected to store the paper on anything other than the author's personal laptop. Upon the sudden and unexpected death of this laptop's CPU on the second to last day of finals week, we were left with no SYE paper and condemned to rewriting this paper over the summer. As of writing this footnote, it is June 28th.}

\begin{figure}
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{png/centralized-vcs.png}
            \caption{
                \centering 
                A centralized version control system requires developers to submit individual file changes to a shared repository containing the software source code.
            }
            \label{fig:centralized-vcs}
        \end{center}
    \end{subfigure}
    \hspace{0.05\textwidth} 
    \begin{subfigure}[c][][c]{0.45\textwidth}
        \begin{center}
            \includegraphics[width=\textwidth]{png/distributed-vcs.png}
            \caption{
                \centering 
                A distributed version control system allows developers to share revised source code repositories directly with eachother, or through a shared server.
            }
            \label{fig:distributed-vcs}
        \end{center}
    \end{subfigure}
    \caption{
        \centering
        Two different types of version control systems for storing and maintaining software code repositories.
    }
    \label{fig:version-control-systems}
\end{figure}

Distributed version control systems store source code history differently than centralized version control systems. Instead of maintaining a single code repository for a software, distributed version control systems allow developers to maintain local copies of a code repository, as shown in figure \ref{fig:distributed-vcs}. When a developer wants to make changes to the repository, they copy the entire repository to their machine, make changes to the relevant files, and share the revised repository with the rest of their team members, either directly or through a shared server.

Distributed version control systems allow developers to modify and test a piece of software without connecting to internet. Additionally, the redundancy of maintaining multiple copies of a code repository ensures that the repository is not lost when a single machine or server fails.

Of course, distributed version control systems have their downsides as well. The architecture is a bit less intuitive than a centralized version control system, so it may be more confusing to understand. Additionally, with multiple copies and version of a code repository existing on different machines, it may be difficult to track down the most up to date version of a software without proper organization. Further, when multiple developers modify code repositories in different ways, there is a potential for versions to conflict. This then requires methods to resolve these conflicts, which can prove to be rather complicated.

\subsubsection{The Git Version Control System}

The git version control system was developed in 2005 to aid in the development of the Linux kernel \footnote{Much of the following information comes from Pro Git, by Scott Chacon and Ben Straub. It is available online for free and is a great source for further investigation into git: \url{https://git-scm.com/book/en/v2}}\cite{ChaconStraub_2014}. From its inception, git was meant to be a free, open-source alternative to commercial version control systems. Today, git has become one of the most popular version control systems on the planet, seeing use by over 90\% of all professional and hobby software developers, according to Stack Overflow's 2022 report \cite{StackOverflow_2022}.

Apart from being freely accessible to the public, git has a few features that have helped its rise in popularity. First, as a distributed version control system, developers have access to an entire code repository on their local development environments. This allows developers to make and test code changes when not connected to the internet. This also makes git actions feel lightning fast, as there is no communication to an external server required for most actions.

Many version control systems provide some support for branches during development. A branch of a code repository is a software version that diverges and can be developed in parallel to a main version. For example, the stable version of a software is often contained on a branch called \lstinline{main}. When developing a new feature to add to this software, a feature can be developed on a branch called \lstinline{feature}, which diverges from \lstinline{main}. Development on \lstinline{feature} can be done safely and separately from \lstinline{main}, and changes can be applied (or \emph{merged}) into \lstinline{main} once the feature is complete.

Git provides support for branching development. Unlike most version control systems though, git branching operations are super quick. Developers can create and switch between branches within seconds, encouraging developers to take full advantage branches during development.

\subsubsection{Git in Snail Language Support}

We use git as a version control system for Snail Language Support. We use git features in a few key ways throughout the development of Snail Language Support.

First, we store the Snail Language Support git repository on GitHub.\footnote{As previously mentioned, the \LaTeX source code for this paper is also hosted on GitHub.} GitHub is a website that allows developers to host their git repositories. As of 2018, GitHub was home to 100 million git repositories all over the world \cite{Warner_2018}. This makes it easy to share changes made to Snail Language Support when asking for advice on how to pursue further development.

We benefitted greatly by using branches during Snail Language Support development. Once a semi-functional extension was developed, further feature and bug fixing development took place in branches. This helped isolate different development ideas, which aided during debugging when code was not performing as expected. These branches were shared on GitHub.

We also took advantage of GitHub pull requests and code reviews when developing Snail Language Support. A pull request is a way to merge a \lstinline{feature} branch back into a \lstinline{main} branch. It allows a user without direct write access to modify a repository by submitting changes and asking the repository owner to accept those changes. Often, before accepting the changes laid out in a pull request, another developer will review the changes made in the code in a code review. A code reviewer can choose to approve the changes outright or ask for further modifications. Once a code review is approved, the pull request is typically merged back into the \lstinline{main} branch of the repository.

We also used GitHub's issue tracking system when developing Snail Language Support. GitHub issues allows developers and users of a git repository to note certain problems with a repository. This might include ideas for future features or documentation of bug fixes. For Snail Language Support, this allowed us to create easily accessible documentation of problems and potential fixes. By documenting these issues on GitHub, they are also easily accessible for future developers of the repository.

\subsection{Software Documentation}

Software documentation is a useful, but underappreciated aspect of the software development role. Software documentation is an essential piece of any software product. This section now describes software documentation in more detail, as well as touching on how documentation is used in Snail Language Support.

\subsubsection{What is Software Documentation?}

Software documentation is any sort of recorded, accessible information that describes a piece of software. This might be a user manual that a customer recieves along with a software product, or a file outlining how other developers might be able to contribute to an open source project.

Most software documentation can be split between two types: internal vs. external. Internal software documentation is created by developers, for developers. This documentation is only accessible to developers of a software, and might be more complicated in nature. They might record known bugs in the software, or might outline the architecture of a service to help developers understand the details of how a service operates.

External software is meant for non-developers, such as consumers of a software product. This documentation is likely written in less technical language, and is meant to give a user an overarching idea of a software, without revealing too many internal details. This might come in the form of a user manual or video guides.

Both internal and external software documentation can take many forms. Software documentation might come in the form of written text in collections of files and manuals. Documentation might also include sample images of user interfaces or diagrams that outline software interactions. Many tools exist to help developers produce useful documentation.

\subsubsection{Documentation in Snail Language Support}

Using documentation formed a huge role in the development of the Snail Language Support extension. We used the VS Code documentation for starter guides on how to build extensions, as well as to look at a few pre-built, example extensions. In addition to big-picture ideas, documentation also contained specific details on how to use the VS Code extension API. We used similar documentation for the language server protocol (LSP) and debug adapter protocol (DAP). Some of the LSP and DAP documentation was hosted on VS Code's documentation website, but most is hosted on separate, LSP-specific and DAP-specific websites.

During development, We also made use of git documentation to look up specific command syntax. Some of this documentation is available while using git command line tools by passing the \lstinline{--help} flag.

During development, we also wanted to focus on creating documentation for the Snail Language Support extension. To do this, Snail Language Support leverages a few different documentation tools: Markdown, Mermaid, and draw.io.

Markdown is a plain-text syntax that takes lightly marked-up text and converts it to clean HTML \cite{Gruber_2004}. The syntax is designed to be lightweight, allowing Markdown text files to be fully legible without converting to HTML. Since it's birth in 2004, Markdown has seen widespread, mainstream adoption across the software industry. Popular tools like GitHub and VS Code offer built-in support for Markdown syntax \cite{GitHub_2023, Microsoft_2023k}. Markdown syntax has also strongly influenced the R Markdown syntax for R programming \cite{RStudio_2020}. Markdown files end with the \lstinline{.md} file extension.

We use Markdown to create a \lstinline{README.md} file which describes valuable information about the Snail Language Support extension for users and developers. We also used Markdown files as notepads to write down thoughts and discoveries during development. These journals helped take note of significant discoveries and learnings during the development process. While these journals are far from polished, most are still stored in the Snail Language Support git repository, in the hopes they may prove useful during future development.

In presenting the Snail Language Support extension (in posters, presentations, and this paper), we found the Mermaid diagramming tool quite helpful. Mermaid is an open source tool that allows developers to create high-quality diagrams using Markdown-inspired syntax \cite{MermaidJS_2023a}. Mermaid diagrams can be developed in Mermaid's online editor, using third-party plug-ins, or by using Mermaid's API in JavaScript files \cite{MermaidJS_2023b}.

We use the third-party Markdown Preview Mermaid Support extension for VS Code to develop Mermaid diagrams in Markdown files \cite{Bierner_2023}. From there, we use the official mermaid command line tool \lstinline{mermaid-cli} to convert the Markdown file to a \lstinline{.png} file \cite{MermaidJS_2023c}. We use this workflow to develop Mermaid sequence diagrams such as those pictured in figure \ref{fig:language-server} and figure \ref{fig:lsp-sequence}.

We also made use of a web-based diagramming tool called draw.io. Draw.io is a free, open source diagramming tool that runs in the web browser \cite{drawio_2023}. Draw.io provides a simple and easy-to-use user interface to allow developers to create high-quality, customisable diagrams quickly. We use the third-party Draw.io Integration extension for VS Code to create and export draw.io diagrams into \lstinline{.png} files \cite{Dieterichs_2023}. This allows us to easily create free-form diagrams such as those seen in figure \ref{fig:directory-structure} and figure \ref{fig:lsp}.

As previously mentioned, creating software documentation \emph{throughout} the Snail Language Support development process was a central goal of this project. Despite this, we found it incredibly difficult to do this in practice. First, it was difficult to motivate ourselves set aside time to create documentation. Instead of documenting steps taken while troubleshooting, we found it easier to save time in the short term by skipping documentation of troubleshooting steps and rushing to further development. Unfortunately, more often than not we would encounter a similar issue later during development, and would have to repeat the same, painstakingly slow troubleshooting process. Trying to save time in the short term by skipping documentation caused us to lose development time in the long run.

We also didn't realize how difficult it is to make high-quality documentation. As frequent users of documentation, we know what lackluster documentation looks like when we see it. However, when trying to make documentation for Snail Language Support, we didn't realize the attention to detail and thoroughness that is required to generate high-quality documentation for a piece of software. We found ourselves getting tired of creating reproducible installation instructions, linking relevant help sites, and trying to ensure instructions were not operating-system-specific. After developing Snail Language Support, we have a newfound appreciation for high-quality documentation.

\section{Conclusions}

This paper presents the Snail Language Support Visual Studio Code (VS Code) extension for the snail programming language. We uncover the motivation behind building this extension, touching on the history of integrated development environments and debugging tools. We discuss the structure of a language support VS Code extension, particularly highlighting the roles of the language server protocol (LSP) to provide advanced error diagnostics and the debug adapter protocol to support future implementation of a snail debugger. We conclude by discussing good version control and documentation practices in software development.

%== BIBLIOGRAPHY ==%
\newpage

\bibliography{refs.bib}
\bibliographystyle{plain}

%== APPENDICES ==%
\newpage

\begin{appendices}

\section{The \lstinline{extension.ts} File}

    The following is the \lstinline{extension.ts} file from Snail Language Support, implemented in TypeScript.

    \lstinputlisting[language=TypeScript]{
        /Users/charlesreinhardt/Git/snail-language-support/client/src/extension.ts
    }
    \label{app:extension-ts}

\newpage
\section{The \lstinline{package.json} File}

    The following is the \lstinline{package.json} file from Snail Language Support.
    \lstinputlisting[language=json]{
        % this absolute path must be different when building on different machines
        /Users/charlesreinhardt/Git/snail-language-support/package.json
    }
    \label{app:package-json}

\newpage
\section{The \lstinline{language-configuration.json} File}

    The following is the \lstinline{language-configuration.json} file from Snail Language Support.
    \lstinputlisting[language=json]{
        /Users/charlesreinhardt/Git/snail-language-support/configurations/language-configuration.json
    }
    \label{app:language-configuration}

\newpage
\section{The \lstinline{snippets.json} File}

    The following is the \lstinline{snippets.json} file from Snail Language Support.
    \lstinputlisting[language=json]{
        /Users/charlesreinhardt/Git/snail-language-support/snippets/snippets.json
    }
    \label{app:snippets-json}

\newpage
\section{TextMate Grammar in Snail Language Support}

    The following is the \lstinline{snail.tmLanguage.json} file from Snail Language Support.
    \lstinputlisting[language=json]{
        /Users/charlesreinhardt/Git/snail-language-support/syntaxes/snail.tmLanguage.json
    }
    \label{app:snail-tmLanguage}

\newpage
\section{The \lstinline{server.ts} File}

    The following is the \lstinline{server.ts} file from Snail Language Support, implemented in TypeScript.
    \lstinputlisting[language=TypeScript]{
        /Users/charlesreinhardt/Git/snail-language-support/client/src/server.ts
    }
    \label{app:server-ts}

\newpage
\section{The \lstinline{debugAdapter.ts} File}

    The following is the \lstinline{debugAdapter.ts} file, which defines the extension-level debug adapter for Snail Language Support.
    \lstinputlisting[language=TypeScript]{
        /Users/charlesreinhardt/Git/snail-language-support/client/src/debugAdapter.ts
    }
    \label{app:debug-adapter-ts}

\newpage
\section{The \lstinline{snaildap.re} File}

    The following is the \lstinline{snaildap.re} file from the snail language interpreter, which defines the snail-level debug adapter used to respond to debug adapter protocol requests.
    \lstinputlisting[]{
        /Users/charlesreinhardt/Git/snail/src/snaildap.re
    }
    \label{app:snaildap}

\newpage
\section{Future Debugger Work}

    The snail-level debug adapter module allows a VS Code client to start and stop a snail debugging session through the VS Code UI. In order to perform debugging-specific actions such as step in/out behavior and variable inspection, a number of DAP requests must be processed and responded to in the \lstinline{snaildap} module. Those requests are listed here, organized by the functionality they are meant to support. They are described in more detail on the Debug Adapter Protocol documentation: \url{https://microsoft.github.io/debug-adapter-protocol/specification}

    \begin{enumerate}
        \item Breakpoints
            \begin{itemize}
                \item BreakpointLocations Request
                \item SetBreakpoints Request
            \end{itemize}
        \item Program execution start/stopping
            \begin{itemize}
                \item Pause Request
                \item Continue Request
            \end{itemize}
        \item Step through
            \begin{itemize}
                \item Next Request
                \item StepInTargets Request
                \item StepIn Request
                \item StepOut Request
            \end{itemize}
        \item Call stack display
            \begin{itemize}
                \item StackTrace Request
            \end{itemize}
        \item Variable/expression inspection
            \begin{itemize}
                \item Scopes Request
                \item Variables Request
                \item Evaluate Request
            \end{itemize}
    \end{enumerate}
    \label{app:further-dap}

\end{appendices}

\end{document}